Efficient ancestry and mutation simulation with msprime 1.0

### 2021-08-13 paper.tex
Abstract pass

\begin{abstract}
Coalescent simulation is a key tool in population genetics and
has been integral to theoretical developments since its beginnings.
Because of the ease with which the basic coalescent model can be simulated,
a large number of different simulators have been developed. However,
the coalescent with recombination is far more challenging to simulate,
and, until recently, it was not possible to simulate efficiently.
The \msprime\ software has revolutionised
coalescent simulation, making it possible to simulate millions
of whole genomes for the first time. We summarise the many features
of \msprime\ 1.0, which is built around a core model of efficiently
implementing recombination via the succinct tree sequence data
structure. Moreover, development of \msprime\ is collaborative
and open source, which reduces duplication of effort and promotes
software quality via community engagement.
% "Bazaar" development model: https://en.wikipedia.org/wiki/The_Cathedral_and_the_Bazaar
\end{abstract}

\textbf{Keywords:} Coalescent simulation, Python, Ancestral Recombination Graphs

>>>>

\begin{abstract}
Stochastic simulation is a key tool in population genetics,
since the models involved are often analytically intractable
and simulation is usually the only way of obtaining ground-truth data to
evaluate inferences.
Because of this necessity, a large number
of specialised simulation programs have been developed, each
filling a particular niche, but with largely overlapping functionality
and a substantial duplication of effort.
Since its introduction, the \msprime\ simulator has substantially
expanded the scope of coalescent simulation, enabling the
simulation and efficient analysis of datasets on a far
larger scale than previously possible.
Here, we introduce \msprime\ version 1.0,
which is built around a core model of efficiently
implementing ancestry and mutation simulations using
the succinct tree sequence data structure and \tskit\ library.
We summarise \msprime's many features, and show that
its performance is excellent, often many times faster
and more memory efficient than specialised alternatives.
These high-performance features have been built using a collaborative,
open source development model,
which reduces duplication of effort and promotes
software quality via community engagement.
% "Bazaar" development model: https://en.wikipedia.org/wiki/The_Cathedral_and_the_Bazaar
\end{abstract}

\textbf{Keywords:} Simulation, Coalescent, Mutations, Ancestral Recombination
Graphs
